Competing Bandits: The Perils of Exploration under Competition

### 2018-09-05 paper/content/introduction.tex
more cosmetic fixes
update intro

We ask whether competition incentivizes adoption of "better" algorithms for exploration. We investigate this issue via extensive numerical experiments in a stylized duopoly model. In our model, two firms compete for users, and simultaneously learn from them. Each firm commits to a multi-armed bandit algorithm, and \emph{explores} according to this algorithm. Users select between the two firms based on the current reputation score: rewards from the firm's algorithm, averaged over a recent time window. Each firm's objective is to maximize its market share (the fraction of users choosing this firm).

@@

Our findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy. In this scenario, the incumbent firm enjoys both a ``data advantage" and a ``reputational advantage". We further investigate which of the two is a stronger barrier to entry. We find that both are strong barriers on its own: removing either one substantially helps the late entrant. Further, the data advantage tends to be a stronger barrier when the incumbent commits to a more advanced bandit algorithm.

>>>>

We ask whether competition incentivizes adoption of ``better" algorithms for exploration. We investigate this issue via extensive numerical experiments in a stylized duopoly model. In our model, two firms compete for users, and simultaneously learn from them. Each firm commits to a multi-armed bandit algorithm, and \emph{explores} according to this algorithm. Users select between the two firms based on the current reputation score: rewards from the firm's algorithm, averaged over a recent time window. Each firm's objective is to maximize its market share (the fraction of users choosing this firm).

@@

Our findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy. \gadelete{In this scenario, the incumbent firm enjoys both a ``data advantage" and a ``reputational advantage". We further investigate which of the two is a stronger barrier to entry. We find that both are strong barriers on its own: removing either one substantially helps the late entrant. Further, the data advantage tends to be a stronger barrier when the incumbent commits to a more advanced bandit algorithm.}

\gaedit{Being first in the market gives firms free data to learn from (a ``data advantage") as well a larger reputation compared to an entrant (a ``reputation advantage"). We investigate which of the two is a stronger barrier to entry. We find that both are strong barriers on their own: removing either one still gives the incumbent a substantially larger share of the market compared to the case where both are removed. Further, the data advantage tends to be a stronger barrier when the incumbent commits to a more advanced bandit algorithm. This occurs because the lack of initial competition allows the incumbent to gather data on actions that may be harder to gather in competition (due to reputational consequences) but that are useful for learning and maximizing long-run reputation.
}
